Using Genetic Algorithms to Optimize the Analogue Method for Precipitation Forecasting in the Swiss Alps

### 2015-11-19 Optimization case study - Journal of Hydrology/elsarticle-template.tex
Using nomenclature.

The method based on the analogy of the synoptic circulation consists in the following steps (Table \ref{table:params_R1}): The similarity of the atmospheric circulation of a target date with every day of the archive is assessed by processing the S1 criteria \citep[Eq.\ (\ref{eq:S1}), ][]{Teweles1954, Drosdowsky2003}, which is a comparison of gradients, over a certain spatial window. \citet{Bontron2005} showed that the geopotential height at 500~hPa and 1000~hPa are the best first predictors of the NCEP/NCAR reanalysis dataset, and that the S1 criteria performs better than scores based on absolute distances. The reason for such better results is that the S1 criteria allows comparing the circulation pattern, by means of the gradients, rather than the absolute value of the geopotential height. To cope with seasonal effects, candidate dates are extracted within a period of 4 months centred around the target date, for every year of the archive.

\begin{equation}
\label{eq:S1}
S1=100 \frac {\displaystyle \sum_{i} \vert \Delta\hat{z}_{i} - \Delta z_{i} \vert}
{\displaystyle \sum_{i} max\left\lbrace \vert \Delta\hat{z}_{i} \vert , \vert \Delta z_{i} \vert \right\rbrace }
\end{equation}
where $\Delta \hat{z}_{i}$ is the forecast geopotential height difference between the \textit{i}th pair of adjacent points of the grid describing the target situation, and $\Delta z_{i}$ is the corresponding observed geopotential height difference in the candidate situation. The differences are processed separately in both directions. The smaller the S1 values are, the more similar the pressure fields.

\begin{table}[htbp]
	\caption{Parameters of the reference method on the atmospheric circulation.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccc}
			\hline
			Level & Variable & Hour & Criteria & Nb \\ 
			\hline 
			0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{2}{*}{1} & Z 1000~hPa & 12~h & \multirow{2}{*}{S1} & \multirow{2}{*}{50} \\
			& Z 500~hPa & 24~h & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_R1}
\end{table}

@@

\begin{table}[htbp]
	\caption{Parameters of the reference method with moisture variables.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccc}
			\hline 
			Level & Variable & Hour & Criteria & Nb \\ 
			\hline 
			0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{2}{*}{1} & Z 1000~hPa & 12~h & \multirow{2}{*}{S1} & \multirow{2}{*}{70} \\
			& Z 500~hPa & 24~h & & \\ 
			\hline
			\multirow{2}{*}{2} & TPW * RH 850~hPa & 12~h & \multirow{2}{*}{RMSE} & \multirow{2}{*}{30} \\
			& TPW * RH 850~hPa & 24~h & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_R2}
\end{table}

@@

The method we finally optimized for the analogy of the atmospheric circulation, based on 4 levels of the geopotential, and that we will name z4, is made of the following degrees of freedom:

@@

\subsection{Results for the z4 method}

The resulting optimized parameters for z4 vary from one subregion to another. An example of the detailed parameters is provided for the Swiss Chablais in Table \ref{table:params_GA_z4}. The optimized spatial windows are given for every subregion in Figure \ref{fig:spatial_windows_z4}, and the selected pressure levels in Table~\ref{table:levels_GA_z4}.

\begin{table}[htbp]
	\caption{Parameters of the z4 method (analogy on 4 levels of the atmospheric circulation) optimized for the Chablais subregion. The columns are the following: L = level, V = variable, H = hour, D = domain, W = weight, C = criteria, N = number of analogues.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccccc}
			\hline L & V & H & D & W & C & N \\ 
			\hline 
			0 & \multicolumn{6}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{8}{*}{1} &  \multirow{2}{*}{Z300} & \multirow{2}{*}{12~h} & -2.5 - 15.0 \degree E & \multirow{2}{*}{32\%} & \multirow{8}{*}{S1} & \multirow{8}{*}{28} \\
			& & & 37.5 - 57.5 \degree N & & & \\ 
			& \multirow{2}{*}{Z700} & \multirow{2}{*}{24~h} & 2.5 - 10.0 \degree E & \multirow{2}{*}{22\%} & & \\ 
			& & & 42.5 - 45.0 \degree N & & & \\ 
			& \multirow{2}{*}{Z1000} & \multirow{2}{*}{6~h} & -5.0 - 15.0 \degree E & \multirow{2}{*}{24\%} & & \\ 
			& & & 42.5 - 47.5 \degree N & & & \\ 
			& \multirow{2}{*}{Z1000} & \multirow{2}{*}{30~h} & 2.5 - 15.0 \degree E & \multirow{2}{*}{22\%} & & \\ 
			& & & 40.0 - 50.0 \degree N & & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_GA_z4}
\end{table}

\begin{table}[htbp]
	\caption{Pressure levels ($\sim$) automatically selected for the z4 method at the different subregions.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccccccc}
			\hline ID & 300 & 400 & 500 & 600 & 700 & 850 & 925 & 1000 \\ 
			\hline 
			1  & $\sim$ &   &   &   & $\sim$ &   &   & $\sim \sim$ \\
			2  & $\sim$ &   &   &   & $\sim$ &   &   & $\sim \sim$ \\
			3  & $\sim$ &   &   &   & $\sim$ &   &   & $\sim \sim$ \\
			4  & $\sim$ &   &   &   & $\sim$ &   &   & $\sim \sim$ \\
			5  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			6  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			7  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			8  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			9  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			10 &   &   &   & $\sim$ & $\sim$ &   &   & $\sim \sim$ \\
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:levels_GA_z4}
\end{table}

The resulting CRPSS scores are provided in Table \ref{table:scores} and is in average 35.8 \% in calibration and 35.5 \% in validation. The improvement of the CRPSS score relatively to the reference method on the atmospheric circulation is provided in Table \ref{table:scores_diff} and is in average 15.3 \% in calibration and 9.9 \% in validation. The score was also calculated for three precipitation thresholds: P\(\geq\)1 mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10, P10 being the daily precipitation with a 10~year return period (Table \ref{table:scores_thresholds_z4}). The gain in score increases with the precipitation threshold: in average, respectively for the different thresholds, 13.3\%, 15.4\% and 29.1\% in calibration and 7.9\%, 11.1\% and 34.5\% for the validation period. The optimized method improves thus even more the forecasts for days with significant precipitation than usual days.

\begin{table}[htbp]
	\caption{CRPSS score (\%) of the three optimized methods.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccccc}
			\hline 
			ID & \multicolumn{2}{c}{z4} & \multicolumn{2}{c}{z4-hi2} & \multicolumn{2}{c}{z4-hi4} \\ 
			& calib & valid & calib & valid & calib & valid \\ 
			\hline 
			1 & 40.3 & 38.5 & 45.5 & 43.8 & 45.9 & 43.6 \\
			2 & 37.8 & 34.9 & 42.4 & 40.8 & 42.9 & 41.3 \\
			3 & 36.6 & 33.9 & 41.4 & 39.0 & 41.8 & 39.5 \\
			4 & 31.2 & 30.8 & 36.1 & 36.4 & 36.8 & 37.1 \\
			5 & 34.4 & 36.5 & 38.6 & 41.4 & 39.2 & 41.8 \\
			6 & 37.0 & 39.9 & 41.0 & 44.5 & 41.9 & 44.6 \\
			7 & 30.0 & 30.1 & 33.8 & 34.6 & 34.4 & 34.8 \\
			8 & 37.7 & 40.1 & 40.7 & 43.1 & 41.2 & 43.3 \\
			9 & 33.1 & 30.0 & 38.4 & 36.6 & 38.7 & 36.3 \\
			10 & 40.0 & 40.0 & 42.6 & 42.9 & 43.3 & 43.2 \\
			\hline
			av. & 35.8 & 35.5 & 40.1 & 40.3 & 40.6 & 40.5 \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:scores}
\end{table}

\begin{table}[htbp]
	\caption{Improvement (\%) of the CRPSS of the three methods. For z4, the improvement is relative to the reference method on the atmospheric circulation. For z4-hi2 and z4-hi4, the improvement is relative to the reference method on the circulation and moisture variables.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccccc}
			\hline 
			ID & \multicolumn{2}{c}{z4} & \multicolumn{2}{c}{z4-hi2} & \multicolumn{2}{c}{z4-hi4} \\ 
			& calib & valid & calib & valid & calib & valid \\ 
			\hline 
			1 & 13.6 & 13.0 & 11.9 & 9.9 & 12.8 & 9.6 \\
			2 & 13.3 & 7.5 & 11.2 & 9.4 & 12.5 & 10.8 \\
			3 & 12.2 & 9.0 & 13.2 & 12.0 & 14.1 & 13.5 \\
			4 & 14.2 & 10.5 & 12.4 & 10.6 & 14.5 & 12.5 \\
			5 & 14.5 & 8.8 & 12.5 & 9.1 & 14.1 & 10.0 \\
			6 & 13.3 & 8.6 & 11.5 & 8.2 & 14.1 & 8.4 \\
			7 & 22.8 & 10.6 & 19.7 & 15.7 & 21.7 & 16.3 \\
			8 & 20.7 & 11.9 & 19.2 & 13.5 & 20.8 & 14.0 \\
			9 & 14.4 & 9.5 & 15.8 & 15.2 & 16.7 & 14.1 \\
			10 & 14.2 & 9.4 & 12.9 & 11.2 & 14.6 & 11.8 \\
			\hline
			av. & 15.3 & 9.9 & 14.0 & 11.5 & 15.6 & 12.1 \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:scores_diff}
\end{table}

\begin{table}[htbp]
	\caption{Improvement (\%) of the CRPSS for different precipitations thresholds for the optimized z4 method.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccccc}
			\hline 
			ID & \multicolumn{2}{c}{P\(\geq\)1 mm} & \multicolumn{2}{c}{P\(\geq\)0.1\(\cdot\)P10} & \multicolumn{2}{c}{P\(\geq\)0.5\(\cdot\)P10} \\ 
			& calib & valid & calib & valid & calib & valid \\ 
			\hline 
			1 & 10.2 & 9.4 & 8.5 & 7.9 & 17.0 & 14.2 \\ 
			2 & 9.9 & 3.4 & 10.2 & 7.3 & 19.3 & 13.7 \\ 
			3 & 13.3 & 10.5 & 13.3 & 10.9 & 19.7 & 9.7 \\ 
			4 & 11.0 & 7.4 & 12.9 & 10.0 & 23.2 & 23.8 \\ 
			5 & 8.6 & 4.2 & 10.9 & 6.2 & 25.2 & 23.8 \\ 
			6 & 10.5 & 5.1 & 11.1 & 7.1 & 21.2 & 41.1 \\ 
			7 & 24.3 & 12.4 & 33.1 & 26.0 & 71.2 & 104.3 \\ 
			8 & 19.0 & 12.7 & 26.2 & 19.2 & 39.4 & 34.9 \\ 
			9 & 12.4 & 6.8 & 13.8 & 9.9 & 24.9 & 48.1 \\ 
			10 & 13.6 & 6.8 & 14.4 & 6.9 & 29.9 & 31.5 \\ 
			\hline 
			av. & 13.3 & 7.9 & 15.4 & 11.1 & 29.1 & 34.5 \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:scores_thresholds_z4}
\end{table}

To assess the cross-compatibility of the parameters, we applied those optimized for one subregion to the others. The resulting losses or gains of the CRPSS are displayed in Table \ref{table:crossing_z4_calib} for the calibration period and in Table \ref{table:crossing_z4_valid} for the validation period.

\begin{table}[htbp]
	\caption{Losses or gains (in \%) of the CRPSS by applying the optimized parameters for the series in column to those in line. Method z4, calibration period.}
	\centerline{\includegraphics[width=\linewidth]{figures/table_crossing_z4_calib.pdf}}
	\label{table:crossing_z4_calib}
\end{table}

\begin{table}[htbp]
	\caption{Losses or gains (in \%) of the CRPSS by applying the optimized parameters for the series in column to those in line. Method z4, validation period.}
	\centerline{\includegraphics[width=\linewidth]{figures/table_crossing_z4_valid.pdf}}
	\label{table:crossing_z4_valid}
\end{table}

@@

The selections of the pressure levels for the analogy of circulation show a great homogeneity and is spatially consistent (Table \ref{table:levels_GA_z4}). First of all, the level 1000~hPa is selected twice (the first time at 6 or 12~h, and the second always at 30~h) and the 700~hPa is selected once for every subregion (always at 24~h). The level which varies from one subregion to another is the upper level (always at 12~h), that is 300~hPa for the North-West part of the catchment, 500~hPa for most of the subregions, and 600~hPa for the Conches Valley. Its spatial distribution is however homogeneous. The optimizer thus provided consistent selections of pressure levels and temporal windows, which depicts a significant preference in the AM, and the success of our GAs to provide consistent results. The automatic selection of the pressure levels is a big advantage of a global optimization, as it is not possible to assess all combinations of 4 levels with the sequential calibration procedure.

The resulting spatial windows (Figure \ref{fig:spatial_windows_z4}) may look very diverse first, but there are significant similarities for subregions located within the same vicinity. The first 4 subregions are characterized by a large spatial window on the upper level, whereas it becomes smaller for the other subregions. For most subregions, the second level (700~hPa) is represented by thin and longitudinally extended domains. The third level (1000~hPa at 6 or 12~h) also contains longitudinally extended domains, but a bit larger. The last one (1000~hPa at 30~h) is rather large and squared. Surprisingly, subregions number 5 (Left side valleys) and 6 (Southern ridges) have exactly the same spatial windows, which suggest that they behave in a similar way and thus could have been merged. This similarity is a good sign for the accuracy of our optimized parameters.

The scores show significant improvements for both the calibration and the validation period (Table \ref{table:scores_diff}). Even more interestingly, the results for higher precipitation thresholds (Table \ref{table:scores_thresholds_z4}) show the largest improvements. This is of particular interest in the framework of flood forecasting.

The analysis of the parameters cross-compatibility shows that obviously, the parameters are optimal on the calibration period when they are calibrated for the given subregion (Table \ref{table:crossing_z4_calib}). However, the losses in CRPSS are not of the same amplitude between the different subregions. Indeed, the Upper Rhone Valley (7) and moreover the Southeast ridges (8) seem to behave significantly differently. This is indeed a climatological reality, as this subregion is the most sensitive to southerly flows \citep{Horton2012}. Parameter sets from other subregions perform poorly for these subregions, and vice versa. Nevertheless, both subregions seem quite inter-compatible. This points to the importance of taking into account leading meteorological influences during subregion creation, which is not always best represented by the geographical distance separating them. Globally, the same pattern can be observed for the validation period (Table \ref{table:crossing_z4_valid}), but in this case, minor improvements may be observed when crossing the parameters. The reason being that the independant validation period contains other events that may be better forecasted by other parameter sets. The relative small differences in score between parametrizations indicate that even though the parameters may differ significantly, the performance may not be drastically affected. Even a change in the pressure level does not mean a radical drop of the score. A different parametrization leads to a different selection of the analogue days, and thus to an improvement of the forecast of certain weather conditions at the expense of others. This raises the question of optimizing more specifically for extreme weather conditions only.

@@

Two methods were assessed: one with 2 moisture predictors (moisture index on 2 pressure levels or at 2 different hours) named z4-hi2, and one with 4 moisture predictors named z4-hi4. When we introduce 2 predictors for the moisture analogy, the number of degrees of freedom raises to 42, and to 54 when we add 4 predictors.

@@

\subsection{Results for z4-hi2 and z4-hi4 methods}

Due to significant similarities between the results from z4-hi2 and z4-hi4, the latter will only be partly shown in order to improve readability.

As previously, the optimized parameters differ from one subregion to another, and this event to a greater extent. Detailed examples are still provided for the Swiss Chablais subregion in Table \ref{table:params_GA_z4_hi2} for z4-hi2. The resulting spatial windows are displayed in Figure \ref{fig:spatial_windows_z4-hi2} for z4-hi2, along with the selected pressure levels for both the circulation and the moisture analogy (Table \ref{table:levels_GA_z4_hi2}).

>>>>

The method based on the analogy of the synoptic circulation consists in the following steps (Table \ref{table:params_R1}): the similarity of the atmospheric circulation of a target date with every day of the archive is assessed by processing the S1 criteria \citep[Eq.\ (\ref{eq:S1}), ][]{Teweles1954, Drosdowsky2003}, which is a comparison of gradients, over a certain spatial window. \citet{Bontron2005} showed that the geopotential height at 500~hPa and 1000~hPa are the best first predictors of the NCEP/NCAR reanalysis dataset, and that the S1 criteria performs better than scores based on absolute distances. The reason for such better results is that the S1 criteria allows comparing the circulation pattern, by means of the gradients, rather than the absolute value of the geopotential height. To cope with seasonal effects, candidate dates are extracted within a period of 4 months centred around the target date, for every year of the archive. Following the nomenclature proposed by \citet{Horton2016}, this method will be named 2Z.

\begin{equation}
\label{eq:S1}
S1=100 \frac {\displaystyle \sum_{i} \vert \Delta\hat{z}_{i} - \Delta z_{i} \vert}
{\displaystyle \sum_{i} max\left\lbrace \vert \Delta\hat{z}_{i} \vert , \vert \Delta z_{i} \vert \right\rbrace }
\end{equation}
where $\Delta \hat{z}_{i}$ is the forecast geopotential height difference between the \textit{i}th pair of adjacent points of the gridded data describing the target situation, and $\Delta z_{i}$ is the corresponding observed geopotential height difference in the candidate situation. The differences are processed separately in both directions. The smaller the S1 values are, the more similar the pressure fields.

\begin{table}[htbp]
	\caption{Parameters of the reference method on the atmospheric circulation (2Z).}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccc}
			\hline
			Level & Variable & Hour & Criteria & Nb \\ 
			\hline 
			0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{2}{*}{1} & Z 1000~hPa & 12~h & \multirow{2}{*}{S1} & \multirow{2}{*}{50} \\
			& Z 500~hPa & 24~h & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_R1}
\end{table}

@@

\begin{table}[htbp]
	\caption{Parameters of the reference method with moisture variables (2Z-2MI).}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccc}
			\hline 
			Level & Variable & Hour & Criteria & Nb \\ 
			\hline 
			0 & \multicolumn{4}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{2}{*}{1} & Z 1000~hPa & 12~h & \multirow{2}{*}{S1} & \multirow{2}{*}{70} \\
			& Z 500~hPa & 24~h & & \\ 
			\hline
			\multirow{2}{*}{2} & TPW * RH 850~hPa & 12~h & \multirow{2}{*}{RMSE} & \multirow{2}{*}{30} \\
			& TPW * RH 850~hPa & 24~h & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_R2}
\end{table}

@@

The method we finally optimized for the analogy of the atmospheric circulation, based on 4 levels of the geopotential, and that we will name 4Zo, is made of the following degrees of freedom:

@@

\subsection{Results for the 4Zo method}

The resulting optimized parameters for 4Zo vary from one subregion to another. An example of the detailed parameters is provided for the Swiss Chablais in Table \ref{table:params_GA_4Zo}. The optimized spatial windows are given for every subregion in Figure \ref{fig:spatial_windows_4Zo}, and the selected pressure levels in Table~\ref{table:levels_GA_4Zo}.

\begin{table}[htbp]
	\caption{Parameters of the 4Zo method (analogy on 4 levels of the atmospheric circulation) optimized for the Chablais subregion. The columns are the following: L = level, V = variable, H = hour, D = domain, W = weight, C = criteria, N = number of analogues.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccccc}
			\hline L & V & H & D & W & C & N \\ 
			\hline 
			0 & \multicolumn{6}{l}{$\pm 60$ days around the target date} \\
			\hline 
			\multirow{8}{*}{1} &  \multirow{2}{*}{Z300} & \multirow{2}{*}{12~h} & -2.5 - 15.0 \degree E & \multirow{2}{*}{32\%} & \multirow{8}{*}{S1} & \multirow{8}{*}{28} \\
			& & & 37.5 - 57.5 \degree N & & & \\ 
			& \multirow{2}{*}{Z700} & \multirow{2}{*}{24~h} & 2.5 - 10.0 \degree E & \multirow{2}{*}{22\%} & & \\ 
			& & & 42.5 - 45.0 \degree N & & & \\ 
			& \multirow{2}{*}{Z1000} & \multirow{2}{*}{6~h} & -5.0 - 15.0 \degree E & \multirow{2}{*}{24\%} & & \\ 
			& & & 42.5 - 47.5 \degree N & & & \\ 
			& \multirow{2}{*}{Z1000} & \multirow{2}{*}{30~h} & 2.5 - 15.0 \degree E & \multirow{2}{*}{22\%} & & \\ 
			& & & 40.0 - 50.0 \degree N & & & \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:params_GA_4Zo}
\end{table}

\begin{table}[htbp]
	\caption{Pressure levels ($\sim$) automatically selected for the 4Zo method at the different subregions.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccccccc}
			\hline ID & 300 & 400 & 500 & 600 & 700 & 850 & 925 & 1000 \\ 
			\hline 
			1  & $\sim$ &   &   &   & $\sim$ &   &   & $\sim \sim$ \\
			2  & $\sim$ &   &   &   & $\sim$ &   &   & $\sim \sim$ \\
			3  & $\sim$ &   &   &   & $\sim$ &   &   & $\sim \sim$ \\
			4  & $\sim$ &   &   &   & $\sim$ &   &   & $\sim \sim$ \\
			5  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			6  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			7  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			8  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			9  &   &   & $\sim$ &   & $\sim$ &   &   & $\sim \sim$ \\
			10 &   &   &   & $\sim$ & $\sim$ &   &   & $\sim \sim$ \\
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:levels_GA_4Zo}
\end{table}

The resulting CRPSS scores are provided in Table \ref{table:scores} and is in average 35.8 \% in calibration and 35.5 \% in validation. The improvement of the CRPSS score relatively to the reference method on the atmospheric circulation is provided in Table \ref{table:scores_diff} and is in average 15.3 \% in calibration and 9.9 \% in validation. The score was also calculated for three precipitation thresholds: P\(\geq\)1 mm, P\(\geq\)0.1\(\cdot\)P10 and P\(\geq\)0.5\(\cdot\)P10, P10 being the daily precipitation with a 10~year return period (Table \ref{table:scores_thresholds_4Zo}). The gain in score increases with the precipitation threshold: in average, respectively for the different thresholds, 13.3\%, 15.4\% and 29.1\% in calibration and 7.9\%, 11.1\% and 34.5\% for the validation period. The optimized method improves thus even more the forecasts for days with significant precipitation than usual days.

\begin{table}[htbp]
	\caption{CRPSS score (\%) of the three optimized methods.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccccc}
			\hline 
			ID & \multicolumn{2}{c}{4Zo} & \multicolumn{2}{c}{4Zo-2MIo} & \multicolumn{2}{c}{4Zo-4MIo} \\ 
			& calib & valid & calib & valid & calib & valid \\ 
			\hline 
			1 & 40.3 & 38.5 & 45.5 & 43.8 & 45.9 & 43.6 \\
			2 & 37.8 & 34.9 & 42.4 & 40.8 & 42.9 & 41.3 \\
			3 & 36.6 & 33.9 & 41.4 & 39.0 & 41.8 & 39.5 \\
			4 & 31.2 & 30.8 & 36.1 & 36.4 & 36.8 & 37.1 \\
			5 & 34.4 & 36.5 & 38.6 & 41.4 & 39.2 & 41.8 \\
			6 & 37.0 & 39.9 & 41.0 & 44.5 & 41.9 & 44.6 \\
			7 & 30.0 & 30.1 & 33.8 & 34.6 & 34.4 & 34.8 \\
			8 & 37.7 & 40.1 & 40.7 & 43.1 & 41.2 & 43.3 \\
			9 & 33.1 & 30.0 & 38.4 & 36.6 & 38.7 & 36.3 \\
			10 & 40.0 & 40.0 & 42.6 & 42.9 & 43.3 & 43.2 \\
			\hline
			av. & 35.8 & 35.5 & 40.1 & 40.3 & 40.6 & 40.5 \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:scores}
\end{table}

\begin{table}[htbp]
	\caption{Improvement (\%) of the CRPSS of the three methods. For 4Zo, the improvement is relative to the reference method on the atmospheric circulation. For 4Zo-2MIo and 4Zo-4MIo, the improvement is relative to the reference method on the circulation and moisture variables.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccccc}
			\hline 
			ID & \multicolumn{2}{c}{4Zo} & \multicolumn{2}{c}{4Zo-2MIo} & \multicolumn{2}{c}{4Zo-4MIo} \\ 
			& calib & valid & calib & valid & calib & valid \\ 
			\hline 
			1 & 13.6 & 13.0 & 11.9 & 9.9 & 12.8 & 9.6 \\
			2 & 13.3 & 7.5 & 11.2 & 9.4 & 12.5 & 10.8 \\
			3 & 12.2 & 9.0 & 13.2 & 12.0 & 14.1 & 13.5 \\
			4 & 14.2 & 10.5 & 12.4 & 10.6 & 14.5 & 12.5 \\
			5 & 14.5 & 8.8 & 12.5 & 9.1 & 14.1 & 10.0 \\
			6 & 13.3 & 8.6 & 11.5 & 8.2 & 14.1 & 8.4 \\
			7 & 22.8 & 10.6 & 19.7 & 15.7 & 21.7 & 16.3 \\
			8 & 20.7 & 11.9 & 19.2 & 13.5 & 20.8 & 14.0 \\
			9 & 14.4 & 9.5 & 15.8 & 15.2 & 16.7 & 14.1 \\
			10 & 14.2 & 9.4 & 12.9 & 11.2 & 14.6 & 11.8 \\
			\hline
			av. & 15.3 & 9.9 & 14.0 & 11.5 & 15.6 & 12.1 \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:scores_diff}
\end{table}

\begin{table}[htbp]
	\caption{Improvement (\%) of the CRPSS for different precipitations thresholds for the optimized 4Zo method.}
	\footnotesize
	\begin{center}
		\begin{tabular}{ccccccc}
			\hline 
			ID & \multicolumn{2}{c}{P\(\geq\)1 mm} & \multicolumn{2}{c}{P\(\geq\)0.1\(\cdot\)P10} & \multicolumn{2}{c}{P\(\geq\)0.5\(\cdot\)P10} \\ 
			& calib & valid & calib & valid & calib & valid \\ 
			\hline 
			1 & 10.2 & 9.4 & 8.5 & 7.9 & 17.0 & 14.2 \\ 
			2 & 9.9 & 3.4 & 10.2 & 7.3 & 19.3 & 13.7 \\ 
			3 & 13.3 & 10.5 & 13.3 & 10.9 & 19.7 & 9.7 \\ 
			4 & 11.0 & 7.4 & 12.9 & 10.0 & 23.2 & 23.8 \\ 
			5 & 8.6 & 4.2 & 10.9 & 6.2 & 25.2 & 23.8 \\ 
			6 & 10.5 & 5.1 & 11.1 & 7.1 & 21.2 & 41.1 \\ 
			7 & 24.3 & 12.4 & 33.1 & 26.0 & 71.2 & 104.3 \\ 
			8 & 19.0 & 12.7 & 26.2 & 19.2 & 39.4 & 34.9 \\ 
			9 & 12.4 & 6.8 & 13.8 & 9.9 & 24.9 & 48.1 \\ 
			10 & 13.6 & 6.8 & 14.4 & 6.9 & 29.9 & 31.5 \\ 
			\hline 
			av. & 13.3 & 7.9 & 15.4 & 11.1 & 29.1 & 34.5 \\ 
			\hline 
		\end{tabular} 
	\end{center}
	\label{table:scores_thresholds_4Zo}
\end{table}

To assess the cross-compatibility of the parameters, we applied those optimized for one subregion to the others. The resulting losses or gains of the CRPSS are displayed in Table \ref{table:crossing_4Zo_calib} for the calibration period and in Table \ref{table:crossing_4Zo_valid} for the validation period.

\begin{table}[htbp]
	\caption{Losses or gains (in \%) of the CRPSS by applying the optimized parameters for the series in column to those in line. Method 4Zo, calibration period.}
	\centerline{\includegraphics[width=\linewidth]{figures/table_crossing_z4_calib.pdf}}
	\label{table:crossing_4Zo_calib}
\end{table}

\begin{table}[htbp]
	\caption{Losses or gains (in \%) of the CRPSS by applying the optimized parameters for the series in column to those in line. Method 4Zo, validation period.}
	\centerline{\includegraphics[width=\linewidth]{figures/table_crossing_z4_valid.pdf}}
	\label{table:crossing_4Zo_valid}
\end{table}

@@

The selections of the pressure levels for the analogy of circulation show a great homogeneity and is spatially consistent (Table \ref{table:levels_GA_4Zo}). First of all, the level 1000~hPa is selected twice (the first time at 6 or 12~h, and the second always at 30~h) and the 700~hPa is selected once for every subregion (always at 24~h). The level which varies from one subregion to another is the upper level (always at 12~h), that is 300~hPa for the North-West part of the catchment, 500~hPa for most of the subregions, and 600~hPa for the Conches Valley. Its spatial distribution is however homogeneous. The optimizer thus provided consistent selections of pressure levels and temporal windows, which depicts a significant preference in the AM, and the success of our GAs to provide consistent results. The automatic selection of the pressure levels is a big advantage of a global optimization, as it is not possible to assess all combinations of 4 levels with the sequential calibration procedure.

The resulting spatial windows (Figure \ref{fig:spatial_windows_4Zo}) may look very diverse first, but there are significant similarities for subregions located within the same vicinity. The first 4 subregions are characterized by a large spatial window on the upper level, whereas it becomes smaller for the other subregions. For most subregions, the second level (700~hPa) is represented by thin and longitudinally extended domains. The third level (1000~hPa at 6 or 12~h) also contains longitudinally extended domains, but a bit larger. The last one (1000~hPa at 30~h) is rather large and squared. Surprisingly, subregions number 5 (Left side valleys) and 6 (Southern ridges) have exactly the same spatial windows, which suggest that they behave in a similar way and thus could have been merged. This similarity is a good sign for the accuracy of our optimized parameters.

The scores show significant improvements for both the calibration and the validation period (Table \ref{table:scores_diff}). Even more interestingly, the results for higher precipitation thresholds (Table \ref{table:scores_thresholds_4Zo}) show the largest improvements. This is of particular interest in the framework of flood forecasting.

The analysis of the parameters cross-compatibility shows that obviously, the parameters are optimal on the calibration period when they are calibrated for the given subregion (Table \ref{table:crossing_4Zo_calib}). However, the losses in CRPSS are not of the same amplitude between the different subregions. Indeed, the Upper Rhone Valley (7) and moreover the Southeast ridges (8) seem to behave significantly differently. This is indeed a climatological reality, as this subregion is the most sensitive to southerly flows \citep{Horton2012}. Parameter sets from other subregions perform poorly for these subregions, and vice versa. Nevertheless, both subregions seem quite inter-compatible. This points to the importance of taking into account leading meteorological influences during subregion creation, which is not always best represented by the geographical distance separating them. Globally, the same pattern can be observed for the validation period (Table \ref{table:crossing_4Zo_valid}), but in this case, minor improvements may be observed when crossing the parameters. The reason being that the independant validation period contains other events that may be better forecasted by other parameter sets. The relative small differences in score between parametrizations indicate that even though the parameters may differ significantly, the performance may not be drastically affected. Even a change in the pressure level does not mean a radical drop of the score. A different parametrization leads to a different selection of the analogue days, and thus to an improvement of the forecast of certain weather conditions at the expense of others. This raises the question of optimizing more specifically for extreme weather conditions only.

@@

Two methods were assessed: one with 2 moisture predictors (moisture index on 2 pressure levels or at 2 different hours) named 4Zo-2MIo, and one with 4 moisture predictors named 4Zo-4MIo. When we introduce 2 predictors for the moisture analogy, the number of degrees of freedom raises to 42, and to 54 when we add 4 predictors.

@@

\subsection{Results for 4Zo-2MIo and 4Zo-4MIo methods}

Due to significant similarities between the results from 4Zo-2MIo and 4Zo-4MIo, the latter will only be partly shown in order to improve readability.

As previously, the optimized parameters differ from one subregion to another, and this event to a greater extent. Detailed examples are still provided for the Swiss Chablais subregion in Table \ref{table:params_GA_4Zo_2MIo} for 4Zo-2MIo. The resulting spatial windows are displayed in Figure \ref{fig:spatial_windows_4Zo-2MIo} for 4Zo-2MIo, along with the selected pressure levels for both the circulation and the moisture analogy (Table \ref{table:levels_GA_4Zo_2MIo}).
